Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2022-04-28 ms/manuscript.Rmd
Update MS
According to corrections made to proof

Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1^, and Atsushi Ebihara^3^

@@

@@

Joel H. Nitta, #601 Iwasaki Lab, Bioscience Bldg, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-0882, Japan

@@

**Premise:** Biodiversity is often only measured with species richness; however, this measure ignores evolutionary history and is not sufficient for making conservation decisions.
Here, we characterize multiple facets and drivers of biodiversity to understand how these relate to bioregions and conservation status in the ferns of Japan.

@@

**Results:** Taxonomic richness was highest at mid latitudes.
Phylogenetic and functional diversity and phylogenetic endemism were highest in small southern islands.
Relative phylogenetic and functional diversity were high at high and low latitudes, and low at mid latitudes.
Grid cells were grouped into `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions.
Temperature and apomixis were identified as drivers of biodiversity patterns.
Conservation status was generally high for grid cells with significantly high biodiversity, but the threat due to herbivory by deer was greater for taxonomic richness than other metrics.

@@

Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]; these sequences originate from samples that were identified using the same taxonomic system as the occurrence data [@Ebihara2010; @Ebihara2019b], so we used these preferentially over other data available on GenBank that might include misidentifications or different taxon concepts.
However, using this data set alone for phylogenetic analysis suffers two drawbacks: it is not densely sampled enough to resolve some internal nodes, and it lacks many lineages that have fossils available for molecular dating.
Furthermore, use of community phylogenies generated by only by sampling the species present in the local community has been shown to produce spurious results in simulation studies [@Park2018].
Because one goal of the study is to understand the distribution of paleo- vs. neoendemic areas (defined in units of time), we require an ultrametric tree.
Therefore, to obtain a robust, ultrametric tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid data set including all fern sequences on GenBank for four widely sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes [FTOL v0.0.1\; @Nitta2022b] (`r ape::Ntip(plastome_tree) %>% number()` operational taxonomic units [OTUs] total, including outgroup taxa).
This gene sampling is comparable to a recent global fern phylogeny that resolved relationships across `r ca` `r number(4000)` taxa using six plastid markers [@Testo2016a], and the addition of plastome data can be expected to increase support along the backbone.

@@

We measured phylogenetic diversity (PD) using Faith's PD [the total branch length connecting all terminal taxa in a community, including the root of the phylogeny\; @Faith1992].
To detect areas with unusually long or short branches, we calculated relative phylogenetic diversity (RPD), which is a ratio of observed PD to PD measured on a tree where all branch lengths have been transformed to equal length [@Mishler2014].

@@

Grid cells with significantly low PD were almost all located on the main islands of Japan (Honshu, Kyushu, and Shikoku; `r figure("rand-diversity")`A).
A smaller number of grid cells had significantly high PD, and these were located mostly on small southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (`r eg`, Kii Peninsula).
Grid cells with significantly low RPD were mostly observed in the southern main islands, particularly in and around Kyushu (`r figure("rand-diversity")`B).
Grid cells with significantly high RPD are located both in the small southern islands and northern Honshu and Hokkaido.
Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity (`r figure("rand-diversity")`C, D).

@@

All models were significant relative to null models without any fixed effects (Appendix S1: `r s_table("lrt-null")`).
The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models.
Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`A).
Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`B--E).
Standard effect size of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`G); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`F).
Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD (Appendix S1: `r s_table("aic")`).

@@

The percentage of protected area in grid cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`A).
Cells with significantly high biodiversity had a similar or greater percentage of area protected compared to the baseline rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
In contrast, grid cells with significantly high biodiversity had a similar or lower percentage of area protected than the baseline rate for Japan at the high protection status (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`) for most biodiversity metrics.
Only grid cells with significantly high PE had a higher percentage area protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total) than the baseline rate at both levels of protection.

@@

The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of the southern subtropical islands (Ryukyu and Ogasawara Islands) (`r figure("rand-diversity")`A, C).
This pattern is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country, which lack subtropical climate; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Island).
The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Appendix S1: Tables `r s_table_num("phy-sig-binary")`, `r s_table_num("phy-sig")`) and suggests that at least in the ferns of Japan, phylogeny can be used as a reasonable stand-in for functional diversity.
Given their high PD, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism (PE); indeed, the vast majority of grid cells with significant levels of PE, as well as paleoendemic grid cells, are located in the southern islands (`r figure("endemism")`; Appendix S1: `r s_figure("endemism-restricted")`).
The high PE observed on the southern subtropical islands is clearly due to the small land area of these islands: small area combined with high PD is expected to lead to high PE.

@@

Water limitation is thought to play an important role in promoting apomixis in ferns especially in deserts [@Grusz2021], since ferns depend on water for transfer of sperm to egg, and apomictic plants would be able to reproduce in the absence of water.
A similar argument has been made to assert that areas with seasonal monsoons are linked to increased rates of apomixis in ferns [@Liu2012; @Tanaka2014; @Picard2021], but in Japan, although the areas that include high percentages of apomictic taxa are monsoonal, the degree of seasonal water limitation does not approach that of a desert.
The monsoon hypothesis was not supported by our results, which showed neither a particularly strong effect of precipitation seasonality in most models (`r figure("coeff")`), nor a correlation with % apomictic taxa (Appendix S1: `r s_table("corr")`).

The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group grid cells with similar biodiversity values (`r figure("div-by-tax-region")`).
Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized ad hoc; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014].
Furthermore, the border between bioregions 1 and 2 approximately corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he proposed based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
The agreement between qualitative analysis of bioregions here with previous ad hoc schemes suggests that ferns are useful as bioindicators in Japan.
Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], they did not play a major role in structuring fern bioregions.
The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits phylogenetic bioregions 2 and 3, the regions have some overlap on Yakushima and Amami Islands (`r figure("bioregions")`B).
The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind [@Tryon1970].
Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic bioregion 3 (mostly the Ryukyu Islands; wetter) from bioregion 4 (mostly the Ogasawara Islands; drier) (Appendix S1: `r s_figure("bioregions-env")`).

@@

First, the data include specimens collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan.
Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (Appendix S1: `r s_figure("abundance")`).
Indeed, one promising avenue of research may be to take advantage of the most densely sampled areas to investigate changes in abundance over time.

@@

Additional supporting information may be found online in the Supporting Information section at the end of the article.

@@

**`r s_table("moran")`**. Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.  
**`r s_table("corr")`**. Results of modified *t*-test for correlation between variables while taking into account spatial position.  
**`r s_table("phy-sig-binary")`**. Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.  
**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan.  
**`r s_table("lrt-null")`**. Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).  
**`r s_table("aic")`**. Model fit as measured with conditional Akaike information criterion (cAIC).  
**`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan.  
**`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.  
**`r s_figure("envir")`**. Reproductive mode and environmental data in 20-km grid cells across Japan.  
**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.  
**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 20-km grid cell in the ferns of Japan.  
**`r s_figure("raw-lat")`**. Raw taxonomic diversity, phylogenetic diversity, and functional diversity of the ferns of Japan plotted by latitude.  
**`r s_figure("raw-biplot")`**. Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.  
**`r s_figure("endemism-restricted")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleoendemism), restricted data set including only taxa endemic to Japan.  
**`r s_figure("k-plot")`**. Selection of *k* for bioregions.  
**`r s_figure("dendrogram")`**. Bioregion dendrograms.  
**`r s_figure("div-by-phy-region")`**. Phylogenetic and morphological diversity of the ferns of Japan by phylogenetic bioregion.    
**`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan.  
**`r s_figure("apo-genus")`**. Rate of apomixis in genera of Japanese ferns with > 1 apomictic taxon.  
**`r s_figure("modelfit-not-ult")`**. Relationship between phylogenetic diversity (PD) and relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).  
**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in taxonomic or phylogenetic bioregions arranged by mean annual temperature and annual precipitation.  
**`r s_figure("protected-map")`**. Maps showing overlap of grid cells with significantly high biodiversity for the ferns of Japan and protected areas.  
**`r s_figure("deer-map")`**. Maps showing overlap of grid cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).

@@

**`r figure("coeff") %>% str_to_upper()`**. Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models for the effect of environment and reproductive mode on biodiversity metrics in the ferns of Japan. (A--E) Environmental models (models including temperature). (F, G) Reproductive models (models including % apomictic taxa). Statistical significance was assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *p*-values indicate probability of no difference between the full and null model. Name of response variable and model type indicated above each subplot. Independent variable with greatest absolute *t*-value in bold for each model.

@@

**`r figure("modelfit") %>% str_to_upper()`**. Relationship between biodiversity metrics and selected predictor variables in the ferns of Japan. (A--E) Environmental (env.) models (models including temperature). Precip., precipitation; Temp., temperature; Temp^2^, temperature squared. (F, G) Reproductive (repro.) models (models including % apomictic taxa). Ribbon shows 95% confidence interval of model. Line fit with the focal predictor variable while averaging over other predictors. For phylogenetic diversity (PD), relative PD (RPD), functional diversity (FD), and relative FD (RFD), the standard effect size (SES) was calculated by comparing raw values to a null distribution of 999 random values, and significance (*p*-value) determined using a two-tailed test (see Materials and Methods).

@@

**`r figure("endemism") %>% str_to_upper()`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleoendemism). neo, grid cells with an overabundance of rare short branches; paleo, grid cells with an overabundance of rare long branches; mixed, significantly endemic grid cells that have neither an overabundance of short nor long branches; super, highly significantly endemic grid cells that have neither an overabundance of short nor long branches. Grid cells without significant endemism in grey. For details, see Materials and Methods.

>>>>

Joel H. Nitta^1,4^, Brent D. Mishler^2^, Wataru Iwasaki^1^, and Atsushi Ebihara^3^

@@

^4^Current affiliation: Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, Chiba, Japan

@@

Joel H. Nitta, #601 Bioscience Bldg, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-0882, Japan

@@

**Premise:** Biodiversity is often only measured with species richness; however, this metric ignores evolutionary history and is not sufficient for making conservation decisions.
Here, we characterize multiple facets and drivers of biodiversity to understand how these relate to bioregions and conservation status in the ferns of Japan.

@@

**Results:** Taxonomic richness was highest at mid-latitudes.
Phylogenetic and functional diversity and phylogenetic endemism were highest in small southern islands.
Relative phylogenetic and functional diversity were high at high and low latitudes, and low at mid-latitudes.
Grid cells were grouped into `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions.
Temperature and apomixis were identified as drivers of biodiversity patterns.
Conservation status was generally high for grid cells with significantly high biodiversity, but the threat due to herbivory by deer was greater for taxonomic richness than other metrics.

@@

Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]; these sequences originate from samples that were identified using the same taxonomic system as the occurrence data [@Ebihara2010; @Ebihara2019b], so we used these preferentially over other data available on GenBank that might include misidentifications or different taxon concepts.
However, using this data set alone for phylogenetic analysis suffers two drawbacks: it is not densely sampled enough to resolve some internal nodes, and it lacks many lineages that have fossils available for molecular dating.
Furthermore, use of community phylogenies generated by sampling only the species present in the local community has been shown to produce spurious results in simulation studies [@Park2018].
Because one goal of the study is to understand the distribution of paleo- vs. neoendemic areas (defined in units of time), we required an ultrametric tree.
Therefore, to obtain a robust, ultrametric tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid data set including all fern sequences on GenBank for four widely sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes [FTOL v0.0.1\; @Nitta2022b] (`r ape::Ntip(plastome_tree) %>% number()` operational taxonomic units [OTUs] total, including outgroup taxa).
This gene sampling is comparable to a recent global fern phylogeny that resolved relationships across `r ca` `r number(4000)` taxa using six plastid markers [@Testo2016a], and the addition of plastome data can be expected to increase support along the backbone.

@@

We measured phylogenetic diversity (PD) using Faith's PD [the total branch length connecting all terminal taxa in a community, including the root of the phylogeny\; @Faith1992].
To detect areas with unusually long or short branches, we calculated relative phylogenetic diversity (RPD), which is the ratio of observed PD to PD measured on a tree where all branch lengths have been transformed to equal length [@Mishler2014].

@@

Grid cells with significantly low PD were almost all located on the main islands of Japan (Honshu, Kyushu, and Shikoku; `r figure("rand-diversity")`A).
A smaller number of grid cells had significantly high PD, and these were located mostly on small southern islands (Ryukyu, Izu, Ogasawara) or coastal areas (`r eg`, Kii Peninsula).
Grid cells with significantly low RPD were mostly observed in the southern main islands, particularly in and around Kyushu (`r figure("rand-diversity")`B).
Grid cells with significantly high RPD are located both in the small southern islands and northern Honshu and Hokkaido.
Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity (`r figure("rand-diversity")`C, D).

@@

All models were significant relative to null models without any fixed effects (Appendix S1: `r s_table("lrt-null")`).
The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models.
Richness showed a hump-shaped relationship with temperature, reaching maximum values at intermediate temperatures (`r figure("modelfit")`A).
Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`B--E).
Standard effect size of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`G); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`F).
Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD (Appendix S1: `r s_table("aic")`).

@@

The percentage of protected area in grid cells with significantly high biodiversity ranges from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`A).
Cells with significantly high biodiversity had a similar or greater percentage of area protected compared to the baseline rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
In contrast, grid cells with significantly high biodiversity had a similar or lower percentage of area protected than the baseline rate for Japan at the high protection status (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`) for most biodiversity metrics.
Only grid cells with significantly high PE had a higher percentage area protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total) than the baseline rate at both levels of protection.

@@

The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of the southern subtropical islands (Ryukyu and Ogasawara Islands) (`r figure("rand-diversity")`A, C).
This pattern is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country, which lack subtropical climate; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima).
The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Appendix S1: Tables `r s_table_num("phy-sig-binary")`, `r s_table_num("phy-sig")`) and suggests that at least in the ferns of Japan, phylogeny can be used as a reasonable stand-in for functional diversity.
Given their high PD, it is perhaps not surprising that the southern islands also host high amounts of phylogenetic endemism (PE); indeed, the vast majority of grid cells with significant levels of PE, as well as paleoendemic grid cells, are located in the southern islands (`r figure("endemism")`; Appendix S1: `r s_figure("endemism-restricted")`).
The high PE observed on the southern islands is clearly due to their small land area: small area combined with high PD is expected to lead to high PE.

@@

Water limitation is thought to play an important role in promoting apomixis in ferns especially in deserts [@Grusz2021], since ferns depend on water for transfer of sperm to egg, and apomictic plants would be able to reproduce in the absence of water.
A similar argument has been made to assert that areas with seasonal monsoons are linked to increased rates of apomixis in ferns [@Liu2012; @Tanaka2014; @Picard2021], but in Japan, although the areas that include high percentages of apomictic taxa are monsoonal, the degree of seasonal water limitation does not approach that of a desert.
The monsoon hypothesis was not supported by our results, which showed neither a particularly strong effect of precipitation seasonality in most models (`r figure("coeff")`), nor a correlation of precipitation seasonality with % apomictic taxa (Appendix S1: `r s_table("corr")`).

The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group grid cells with similar biodiversity values (`r figure("div-by-tax-region")`).
Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized ad hoc; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014].
Furthermore, the border between bioregions 1 and 2 approximately corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
The agreement between qualitative analysis of bioregions here with previous ad hoc schemes suggests that ferns are useful as bioindicators in Japan.
Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], they did not play a major role in structuring fern bioregions.
The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits phylogenetic bioregions 2 and 3, these regions overlap near Yakushima (`r figure("bioregions")`B).
The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind [@Tryon1970].
Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic bioregion 3 (mostly the Ryukyu Islands; wetter) from bioregion 4 (mostly the Ogasawara Islands; drier) (Appendix S1: `r s_figure("bioregions-env")`).

@@

First, the data include specimens collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan.
Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (Appendix S1: `r s_figure("abundance")`A).
Indeed, one promising avenue of research may be to take advantage of the most densely sampled areas to investigate changes in abundance over time.

@@

This article includes online-only Supplemental Data.

@@

**`r s_table("moran")`**. Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.  
**`r s_table("corr")`**. Results of modified *t*-test for correlation between variables while taking into account spatial position.  
**`r s_table("phy-sig-binary")`**. Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.  
**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan.  
**`r s_table("lrt-null")`**. Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).  
**`r s_table("aic")`**. Model fit as measured with conditional Akaike information criterion (cAIC).  
**`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan.  
**`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.  
**`r s_figure("envir")`**. Reproductive mode and environmental data in 20-km grid cells across Japan.  
**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.  
**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 20-km grid cell in the ferns of Japan.  
**`r s_figure("raw-lat")`**. Raw biodiversity metrics of the ferns of Japan plotted by latitude.  
**`r s_figure("raw-biplot")`**. Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.  
**`r s_figure("endemism-restricted")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleoendemism), restricted data set including only taxa endemic to Japan.  
**`r s_figure("k-plot")`**. Selection of *k* for bioregions.  
**`r s_figure("dendrogram")`**. Bioregion dendrograms.  
**`r s_figure("div-by-phy-region")`**. Phylogenetic and morphological diversity of the ferns of Japan by phylogenetic bioregion.    
**`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan.  
**`r s_figure("apo-genus")`**. Rate of apomixis in genera of Japanese ferns with > 1 apomictic taxon.  
**`r s_figure("modelfit-not-ult")`**. Relationship between phylogenetic diversity (PD) and relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).  
**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in taxonomic or phylogenetic bioregions arranged by mean annual temperature and annual precipitation.  
**`r s_figure("protected-map")`**. Maps showing overlap of grid cells with significantly high biodiversity for the ferns of Japan and protected areas.  
**`r s_figure("deer-map")`**. Maps showing overlap of grid cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).

@@

**`r figure("coeff") %>% str_to_upper()`**. Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models for the effect of environment and reproductive mode on biodiversity metrics in the ferns of Japan. (A--E) Environmental (env.) models (models including temperature). (F, G) Reproductive (repro.) models (models including % apomictic taxa). Precip., precipitation; Temp., temperature; SES, standard effect size; PD, phylogenetic diversity; RPD, relative PD; FD, functional diversity; RFD, relative FD. Statistical significance was assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *p*-values indicate probability of no difference between the full and null model. Name of response variable and model type indicated above each subplot. Independent variable with greatest absolute *t*-value in bold for each model.

@@

**`r figure("modelfit") %>% str_to_upper()`**. Relationship between biodiversity metrics and selected predictor variables in the ferns of Japan. (A--E) Environmental models (models including temperature). (F, G) Reproductive models (models including % apomictic taxa). Ribbon shows 95% confidence interval of model. Line fit with the focal predictor variable while averaging over other predictors. For phylogenetic diversity (PD), relative PD (RPD), functional diversity (FD), and relative FD (RFD), the standard effect size (SES) was calculated by comparing raw values to a null distribution of 999 random values, and significance (*p*-value) determined using a two-tailed test (see Materials and Methods).

@@

**`r figure("endemism") %>% str_to_upper()`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleoendemism). neo, grid cells with an overabundance of rare short branches; paleo, grid cells with an overabundance of rare long branches; mixed, significantly endemic grid cells that have neither an overabundance of short nor long branches; super, highly significantly endemic grid cells that have neither an overabundance of short nor long branches. Grid cells without significant endemism in light grey. For details, see Materials and Methods.
